---
title: "Fisheries and climate change"
resource_files:
- "shp/MPA_MX_v1.0_01032021/MPA_MX_c1-0.shx" 
- "shp/MPA_MX_v1.0_01032021/MPA_MX_c1-0.prj" 
- "shp/MPA_MX_v1.0_01032021/MPA_MX_c1-0.dbf" 
output: 
  flexdashboard::flex_dashboard: 
    orientation: row
    vertical_layout: fill
    navbar:
      - { title: "GitHub", href: "https://github.com/CBMC-GCMP/dafishr", align: right, icon: "fa-github" }
    theme: 
      version: 4
      bg: "#121212"
      fg: "#FFFFFF" 
      primary: "#b3b3b3"
      navbar-bg: "#03DAC5"
      base_font: 
        google: Prompt
      heading_font:
        google: Sen
      code_font:
        google: 
          # arguments to sass::font_google() 
          family: JetBrains Mono
          local: false
    source_code: embed 
    runtime: shiny 
    

    
---

About {data-icon="fa-info" data-orientation=rows} 
=======================================================================



### Goals

This project aims to inform the fishing industry on how climate change can affect their catches.

### 
Through a series of Machine Learning models we can predict the fishery yields under distinct climate change scenarios.

### 
Fishing and climate scenario datasets would help us build and improve this interactive dashboard, so any user could visualize the map of fishing yields or obtain certain fishing statistics.



Modeled variables {data-icon="fa-globe" data-navmenu="Menu"}
=======================================================================
        
```{r setup, include=FALSE, results='hide'}

library(flexdashboard)
library(leaflet)
library(mapview)
library(rgdal)
library(sf)
library(tidyverse)
library(ggthemes)
library(lubridate)
library(raster)
library(shiny)
library(openair)
library(shinythemes)
library(ggplot2)
library(rmarkdown)
library(RColorBrewer)
library(janitor)


# Loading forecast and heatwaves data ----------------
phw_slope <- raster::raster("data/rasters/heatwaves/raster_pacific_slope.tif")

phw_p <- raster::raster("data/rasters/heatwaves/raster_pacific_p.tif")

p_sst <- raster::raster("data/rasters/heatwaves/pacific_mean_SST.tif")

yield_p <- raster("data/rasters/average_yield_P.tif")


fh <- raster("data/rasters/fishing_hours_sum_2018_2021.tiff")
crs(fh) <- sp::CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")


# Loading MPA layer -------------------
mpa <- raster::shapefile("shp/MPA_MX_v1.0_01032021/MPA_MX_c1-0.shp", encoding = "UTF-8", verbose = FALSE)


# Loading fisheries predictions ----------------

# t_albacares <- raster("data/rasters/mean_MCP/Thunnus_albacares_avg_MCP.tif")
# t_atlanticus <- raster("data/rasters/mean_MCP/Thunnus_atlanticus_avg_MCP.tif")
# t_obesus <- raster("data/rasters/mean_MCP/Thunnus_obesus_avg_MCP.tif")
# t_orientalis <- raster("data/rasters/mean_MCP/Thunnus_orientalis_avg_MCP.tif")
# t_thynnus <- raster("data/rasters/mean_MCP/Thunnus_thynnus_avg_MCP.tif")


```



```{r}


# # Sea Surface Temperature Labels
# sst_df <- as.data.frame(p_sst, xy=T) %>% 
#   filter(!is.na(pacific_mean_SST)) %>% 
#   # st_as_sf(., coords=c("x", "y"), crs=4326) %>% 
#   rename(value=pacific_mean_SST) %>% 
#            mutate(variable="SST")
# 
# 
# 
# # Marine Heatwaves Labels
# mhw_df <- as.data.frame(phw_slope, xy=T) %>% 
#   filter(!is.na(raster_pacific_slope)) %>% 
#   # st_as_sf(., coords=c("x", "y"), crs=4326) %>% 
#   rename(value=raster_pacific_slope) %>% 
#   mutate(variable="mhw")
# 
# 
# # Fishing hours Labels
# fh_df <- as.data.frame(fh, xy=T) %>% 
#   filter(!is.na(fishing_hours_sum_2018_2021)) %>% 
#   # st_as_sf(., coords=c("x", "y"), crs=4326) %>% 
#   rename(value=fishing_hours_sum_2018_2021) %>% 
#   mutate(variable="fh")
# 
# 
# 
# raster_values <- rbind(sst_df, mhw_df, fh_df)

# Sea Surface Temperature Labels
# sst_df$labels1 <- paste0(
#                       "<b style= 'color:#263270'> Value: </b> ", sst_df$value, "<br/> ") %>%
#         lapply(htmltools::HTML)
# 
# # Marine Heatwaves Labels
# raster_values$labels2 <- paste0("<h3 style= 'background-color:#3d0e27; color: white; text-align: center; font-size: 150%; font-family:Times New Roman'> Pixel Value </h3>",
#                       "<b style= 'color:#263270'> Slope value: </b> ", raster_values$value, "<br/> ") %>%
#         lapply(htmltools::HTML)
# 
# 
# # MPA labels
# mpa$labels3 <- paste0("<h3 style= 'background-color:#3a875a; color: white; text-align: center; font-size: 150%; font-family:Times New Roman'> Marine Protected Area </h3>", 
#                       "<b style= 'color:#263270'> Name: </b> ", mpa$ANP, "<br/> ",
#                       "<b style= 'color:#263270'>Management Plan: </b> ", "<a href =\"", mpa$PM_link, "\", target=\"_blank\"> PDF </a>" ,   "<br/> ",
#                       "<b style= 'color:#263270'>Date: </b> ", mpa$FECHA, "<br/> ",
#                       "<b style= 'color:#263270'> Zone: </b> ", mpa$Categor, "<br/> ",
#                       "<b style= 'color:#263270'> Subzone: </b> ", mpa$Subznfc, "<br/> ",
#                       "<b style= 'color:#263270'> Diving site: </b> ", mpa$Buceo, "<br/> ",
#                       "<b style= 'color:#263270'> Fishing site: </b> ", mpa$Pesca, "<br/> ") %>%
#         lapply(htmltools::HTML)
# 
# # Fishing hours Labels
# raster_values$labels4 <- paste0("<h3 style= 'background-color:#3d0e27; color: white; text-align: center; font-size: 150%; font-family:Times New Roman'> Pixel Value </h3>",
#                       "<b style= 'color:#263270'> Hours: </b> ", raster_values$value, "<br/> ") %>%
#         lapply(htmltools::HTML)




heat <- colorNumeric(c("#0c2c84" , "#ffffff", "#850c0c" ), values(phw_slope),
                    na.color = "transparent")

sst <- colorNumeric(c("#1000c2" ,  "orange", "red" ), values(p_sst),
                    na.color = "transparent")

# t <- colorNumeric(c("#47e0ff", "orange", "red"), values(t_obesus), na.color = "transparent")

h <- colorNumeric(c("darkblue", "orange", "red"), values(fh), na.color = "transparent")

# p <- colorNumeric(c("#000000" , "#808080", "#e3e3e3"), values(phw_p), na.color = "transparent")

yield_bins <- c(0.464, 0.470, 0.475, 0.480, 0.485, 0.490, 0.495 )

yield_pal <- colorBin(c("#1c164d", "#e0c200", "#038a0a"), bins=yield_bins)


leaflet() %>%
        # addProviderTiles(providers$Esri.WorldImagery, group = "Satellite") %>%
  addProviderTiles(providers$Esri.OceanBasemap, group = "Océano") %>%
  addRasterImage(fh, colors=h, opacity = 1, layerId = 3, group="Fishing hours 2021") %>%
  addRasterImage(phw_slope, colors=heat , opacity = 0.8, layerId = 1, group="Marine HWs Slope") %>%
  addRasterImage(p_sst, colors= sst, opacity = 1, layerId = 1, group="SST") %>%
  addRasterImage(yield_p, colors= yield_pal, opacity = 1, layerId = 1, group="Yield Probability") %>%
  # 
  # addCircleMarkers(data=
  # 
  
  # 
  # addCircleMarkers(~x, ~y,
  #   popup = ~labels1, color="white", fillColor = 0, stroke = FALSE, fillOpacity = 0, group="SST") %>%
  # addCircleMarkers(lat=subset(raster_values, variable=="mhw")$y, lng=subset(raster_values, variable=="mhw")$x,
  #   popup = ~labels2, color="white", fillColor = 0, stroke = FALSE, fillOpacity = 0,group="Marine HWs Slope") %>%
  # addCircleMarkers(lat=subset(raster_values, variable=="fh")$y, lng=subset(raster_values, variable=="fh")$x,
  #   popup = ~labels4, color="white", fillColor = 0, stroke = FALSE, fillOpacity = 0,group="Fishing hours 2021") %>%
  # addRasterImage(phw_p, colors=p , opacity = 1, layerId = 1, group="Marine HWs P-values") %>%
  # addRasterImage(t_obesus, colors=t , opacity = 0.8, layerId = 1, group="MCP <i>Thunnus obesus</i> ") %>%
  
        #   addPolygons(
        #         data = mpa,
        #         group = "MPA",
        #         weight = 2,
        #         opacity = 1,
        # 
        #         fillColor = "magma",
        #         fillOpacity = 0.3,
        #         color = "green",
        #         dashArray = "2",
        #         highlightOptions = highlightOptions(
        #                 color = "#73d7ff",
        #                 weight = 2,
        #                 bringToFront = TRUE
        #         )
        #        , popup = ~ labels3,
        # ) %>%

  
  

  addLegend(pal=heat, values=values(phw_slope), title="Marine Heatwaves", group="Marine HWs Slope" ) %>%
  addLegend(pal=sst, values=values(p_sst), title="Sea Surface Temperature", position="bottomleft", group="SST" ) %>%
  # addLegend(pal=t, values=values(t_obesus), title="MCP", position="bottomright" ) %>%
  addLegend(pal=h, values=values(fh), title="Fishing Hours", position="topleft", group="Fishing hours 2021" ) %>%
    addLegend(pal=yield_pal, values=values(yield_p), title=" Average Yield Probability (2021-2030)", position="topleft", group="Yield Probability" ) %>%
  # addLegend(pal=p, values=values(phw_p), title="P-values", position = "bottomright") %>%
  

    
  
  
        addLayersControl(
                
                overlayGroups = c(
                                  "SST",
                                  "Marine HWs Slope",
                                  # "Marine HWs P-values",
                                  # '<p style="font-style: italic;">  <font size="-1">Thunnus albacares</font> </p>MCP', 
                                  "Fishing hours 2021",
                                  "Yield Probability"
                                  # ,"MCP <i>Thunnus obesus</i> "
                                  # ,"MPA"
                                  ),
                options = layersControlOptions(collapsed = TRUE)
        ) %>%
          hideGroup(c("Marine HWs Slope","Fishing hours 2021" , "Yield Probability"
                      # ,"MCP <i>Thunnus obesus</i> "
                      # ,"Marine HWs P-values"
                      # , '<p style="font-style: italic;">  <font size="-1">Thunnus albacares</font> </p>MCP'
                      )) %>%
        
        setView(lng = -118,
                lat = 22,
                zoom = 5) 
```

Forecast {data-icon="fa-chart-line" data-navmenu="Menu"}
=======================================================================

```{r echo=FALSE, results= 'hide'}

#Load fisheries forecasts (hrs)

## Tuna
Tuna <- read.csv("data/fishing_hours_forecast/forecast_tuna.csv") %>%
  mutate(effort_catch= effort_catch)
Tuna$date <- as.Date(Tuna$date)

Sardine <- read.csv("data/fishing_hours_forecast/forecast_sardine.csv")%>%
  mutate(effort_catch= effort_catch/1500)
Sardine$date <- as.Date(Sardine$date)

Shark <- read.csv("data/fishing_hours_forecast/forecast_sharks.csv")%>%
  mutate(effort_catch= effort_catch/1500)
Shark$date <- as.Date(Shark$date)

Shrimp <- read.csv("data/fishing_hours_forecast/forecast_tuna.csv")%>%
  mutate(effort_catch= effort_catch/1500)
Shrimp$date <- as.Date(Shrimp$date)

Forecast <- read.csv("data/forecasts.csv")%>%
  mutate(effort_catch= effort_catch/1500)
Forecast$date <- as.Date(Forecast$date)

```

Column {.sidebar}
-------------------------------------

<font size="+2"> Please specify type of fishery and period of time </font>

<br/>
<br/>
<br/>
<br/>
<br/>

```{r}


## Column {.sidebar key =" value, one per line"}

selectInput("fishery", label = strong("Fishery:"),
            choices = c("Tuna", "Shark", "Shrimp", "Sardine", "Forecast"), selected = "Tuna")


dateRangeInput("date", strong("Date range:"),
               start = "2021-12-31",
               end = "2050-12-31",
               min = "2021-12-31",
               max = "2100-12-31")

# sliderInput("date", label = "Date:",
#             min = "2021-12-31", max = "2100-12-31", value = 1, step = 0.2)




```


Column {}
-------------------------------------

###  <font size="+6"> Fishing hours </font>  

```{r echo=FALSE}
renderPlot({
if(input$fishery=="Tuna"){


  Tuna %>%
    filter(

           date >= as.Date(input$date[1]),
           date <= as.Date(input$date[2])) %>%
    ggplot(aes(x = date, y = effort_catch
               # , color = Users
               )) +
    geom_line(size = 1, alpha=0.8, col="#f7cb72") +
    # geom_smooth( method = "gam", methods.arg = list(family = Gamma(log)), fill="#3700B3",color="#03DAC6"
    #              )+

    labs(x = "Date", y = "Fishing hours", title=paste0(input$fishery," fisheries from ", format(input$date[1], format="%Y"), " to ", format(input$date[2], format="%Y") )) +
    theme_classic()+
    theme(axis.text.y = element_text(size=18, colour ="#FFFFFF" ),
          axis.text.x = element_text(size=17, colour ="#FFFFFF" ),
          axis.title.y = element_text(size=20, face="bold", colour ="#FFFFFF" ),
          axis.title.x= element_text(size=20, face="bold", colour ="#FFFFFF" ),
          plot.title = element_text(size=22, face="bold", hjust=0.5, colour ="#FFFFFF" ),
          plot.background = element_rect(fill="#121212"),
          panel.background = element_rect(fill="#121212"),
          axis.line = element_line(colour = "#FFFFFF"))


} else if(input$fishery=="Shark"){

    Shark %>%
    filter(

           date >= as.Date(input$date[1]),
           date <= as.Date(input$date[2])) %>%
    ggplot(aes(x = date, y = effort_catch
               # , color = Users
               )) +
    geom_line(size = 1, alpha=0.8, col="#f7cb72") +
    # geom_smooth( method = "gam", methods.arg = list(family = Gamma(log)), fill="#3700B3",color="#03DAC6"
    #              )+
    labs(x = "Date", y = "Fishing hours", title=paste0(input$fishery," fisheries from ", format(input$date[1], format="%Y"), " to ", format(input$date[2], format="%Y") )) +
    theme_classic()+
    theme(axis.text.y = element_text(size=18, colour ="#FFFFFF" ),
          axis.text.x = element_text(size=17, colour ="#FFFFFF" ),
          axis.title.y = element_text(size=20, face="bold", colour ="#FFFFFF" ),
          axis.title.x= element_text(size=20, face="bold", colour ="#FFFFFF" ),
          plot.title = element_text(size=22, face="bold", hjust=0.5, colour ="#FFFFFF" ),
          plot.background = element_rect(fill="#121212"),
          panel.background = element_rect(fill="#121212"),
          axis.line = element_line(colour = "#FFFFFF"))

} else if(input$fishery=="Shrimp"){

  Shrimp %>%
    filter(

           date >= as.Date(input$date[1]),
           date <= as.Date(input$date[2])) %>%
    ggplot(aes(x = date, y = effort_catch
               # , color = Users
               )) +
    geom_line(size = 1, alpha=0.8, col="#f7cb72") +
    # geom_smooth( method = "gam", methods.arg = list(family = Gamma(log)), fill="#3700B3",color="#03DAC6"
    #              )+
    labs(x = "Date", y = "Fishing hours",  title=paste0(input$fishery," fisheries from ", format(input$date[1], format="%Y"), " to ", format(input$date[2], format="%Y") )) +
    theme_classic()+
    theme(axis.text.y = element_text(size=18, colour ="#FFFFFF" ),
          axis.text.x = element_text(size=17, colour ="#FFFFFF" ),
          axis.title.y = element_text(size=20, face="bold", colour ="#FFFFFF" ),
          axis.title.x= element_text(size=20, face="bold", colour ="#FFFFFF" ),
          plot.title = element_text(size=22, face="bold", hjust=0.5, colour ="#FFFFFF" ),
          plot.background = element_rect(fill="#121212"),
          panel.background = element_rect(fill="#121212"),
          axis.line = element_line(colour = "#FFFFFF"))

} else if(input$fishery=="Sardine"){

   Sardine %>%
    filter(

           date >= as.Date(input$date[1]),
           date <= as.Date(input$date[2])) %>%
    ggplot(aes(x = date, y = effort_catch
               # , color = Users
               )) +
    geom_line(size = 1, alpha=0.8, col="#f7cb72") +
    # geom_smooth( method = "gam", methods.arg = list(family = Gamma(log)), fill="#3700B3",color="#03DAC6"
    #              )+
    labs(x = "Date", y = "Fishing hours", title=paste0(input$fishery," fisheries from ", format(input$date[1], format="%Y"), " to ", format(input$date[2], format="%Y") )) +
    theme_classic()+
    theme(axis.text.y = element_text(size=18, colour ="#FFFFFF" ),
          axis.text.x = element_text(size=17, colour ="#FFFFFF" ),
          axis.title.y = element_text(size=20, face="bold", colour ="#FFFFFF" ),
          axis.title.x= element_text(size=20, face="bold", colour ="#FFFFFF" ),
          plot.title = element_text(size=22, face="bold", hjust=0.5, colour ="#FFFFFF" ),
          plot.background = element_rect(fill="#121212"),
          panel.background = element_rect(fill="#121212"),
          axis.line = element_line(colour = "#FFFFFF"))

} else{

  Forecast %>%
    filter(

           date >= as.Date(input$date[1]),
           date <= as.Date(input$date[2])) %>%
    ggplot(aes(x = date, y = effort_catch
               # , color = Users
               )) +
    geom_line(size = 1, alpha=0.8, col="#f7cb72") +
    # geom_smooth( method = "gam", methods.arg = list(family = Gamma(log)), fill="#3700B3",color="#03DAC6"
    #              )+
    labs(x = "Date", y = "Fishing hours", title=paste0(input$fishery," from ", format(input$date[1], format="%Y"), " to ", format(input$date[2], format="%Y") )) +
    theme_classic()+
    theme(axis.text.y = element_text(size=18, colour ="#FFFFFF" ),
          axis.text.x = element_text(size=17, colour ="#FFFFFF" ),
          axis.title.y = element_text(size=20, face="bold", colour ="#FFFFFF" ),
          axis.title.x= element_text(size=20, face="bold", colour ="#FFFFFF" ),
          plot.title = element_text(size=22, face="bold", hjust=0.5, colour ="#FFFFFF" ),
          plot.background = element_rect(fill="#121212"),
          panel.background = element_rect(fill="#121212"),
          axis.line = element_line(colour = "#FFFFFF"))
}


})

```



Row
-----------------------------------------------------------------------

### Minimum hours predicted

```{r}
renderValueBox({

  if(input$fishery =="Tuna"){

      min <- round(min(Tuna$effort_catch),0)
  valueBox(min,
           icon = "ion-android-arrow-down",
           color = "#198040")
  } else if(input$fishery=="Shark"){
          min <- round(min(Shark$effort_catch),3)
  valueBox(min,
           icon = "ion-android-arrow-down",
           color = "#198040")
  }else if(input$fishery=="Shrimp"){
          min <- round(min(Shrimp$effort_catch),3)
  valueBox(min,
           icon = "ion-android-arrow-down",
           color = "#198040")
  }else if(input$fishery=="Sardine"){
          min <- round(min(Sardine$effort_catch),3)
  valueBox(min,
           icon = "ion-android-arrow-down",
           color = "#198040")
  }else {
          min <- min(Forecast$effort_catch)
  valueBox(min,
           icon = "ion-android-arrow-down",
           color = "#198040")
  }


})
```


### Maximum hours predicted

```{r}
renderValueBox({

  if(input$fishery =="Tuna"){

      max <- round(max(Tuna$effort_catch),0)
  valueBox(max,
           icon = "ion-android-arrow-up",
           color = "#801919")
  }else if(input$fishery=="Shark"){
          max <- round(max(Shark$effort_catch),3)
  valueBox(max,
           icon = "ion-android-arrow-up",
           color = "#801919")
  }else if(input$fishery=="Shrimp"){
          max <- round(max(Shrimp$effort_catch),3)
  valueBox(max,
           icon = "ion-android-arrow-up",
           color = "#801919")
  }else if(input$fishery=="Sardine"){
          max <- round(max(Sardine$effort_catch),3)
  valueBox(max,
           icon = "ion-android-arrow-up",
           color = "#801919")


   } else {
          max <- max(Forecast$effort_catch)
  valueBox(max,
           icon = "ion-android-arrow-up",
           color = "#801919")}

})
```